Sakura Blossom Prediction Model for Japan

### 2024-12-02 paper/paper.qmd
finish !!!

---
title: "Sakura Blossom Prediction Model for Japan"
subtitle: "Forecasting Sakura Blossom Using Bayesian Hierarchical Regression"
author: 
  - Shanjie Jiao
thanks: "Code and data are available at: https://github.com/Jie-jiao05/Sakura-Blossom-Prediction-Model."
date: today
date-format: long
abstract: "This study develops a Bayesian hierarchical linear regression model to predict sakura blooming dates across Japan by analyzing key predictors such as temperature, latitude, and longitude. The model demonstrates strong predictive accuracy, capturing over 95% of the variability in flowering dates and revealing clear regional and climatic patterns. The findings provide valuable insights into the effects of climate change on sakura phenology, aiding tourism planning and ecological conservation. Future improvements could incorporate dynamic modeling techniques, additional environmental variables, and broader datasets to enhance the model's precision and adaptability."
format: pdf
toc: true
number-sections: true
bibliography: references.bib

@@

library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(lubridate)
library(rvest)
library(palmerpenguins)
library(maps)
library(ggplot2)
library(knitr)
library(arrow)
library(rstanarm)
library(plotly)
library(patchwork)
library(tidyr)
library(bayesplot)
library(gridExtra)

@@

Sakura not merely a ornamental plants but also hold profound cultural significance. In Japanese literature, poetry, and art, sakura blossoms carry deep emotional and symbolic meaning, with the aesthetic concept of "mono no aware" being particularly notable. Due to their short blooming period, sakura blossoms are often seen as a metaphor for the impermanence and fleeting beauty of life, evoking deep reflection and appreciation for the essence of existence.

@@

Given the importance of sakura blooming times for tourism planning and economic activities, accurately forecasting these dates is essential. This study aims to utilize linear regression and Bayesian spline methods to systematically analyze the effects of temperature and geographical location on sakura blooming times. By developing a predictive model, the study seeks to provide scientific insights for sakura enthusiasts worldwide, as well as for tourism and related industries, facilitating more precise planning of viewing activities and resource allocation. Furthermore, analyzing sakura blossom data can also expore on the impact of global warming on blooming periods."

The Bayesian hierarchical model build in this research predicts sakura flowering dates with high accuracy, achieving RMSE values of 3.1 (training) and 3.4 (testing) while explaining over 95% of variability.  Notably, temperature and latitude emerge as significant predictors, revealing earlier blooming in southern regions and highlighting strong sensitivity to climatic and geographic factors. Also as emperature emerges as the most significant predictor, with an increase of 1°C in mean monthly temperature advancing flowering dates by approximately 0.33 days. Latitude shows a strong positive effect, with higher latitudes leading to later flowering dates. In contrast, longitude has a negligible impact.

The structure of this paper is as follows:  @sec-data details the data sources and the methodologies employed, including data scraping and manipulation techniques.  @sec-model outlines the development of prediction models, specifically Linear Regression and Bayesian Spline Models, which are further analyzed in  @sec-result. In  @sec-dis, the impact of global warming on the sakura blossom period, along with real-life implementation and limitations of the study, will be discussed, also how sakura flowering condition from history to present, and will providing insights for further improvement. Further dataset and model detail will be presented in @sec-sam and @sec-model-details separately.

@@

We used the statistical programming language R [@citeR] to perform all analyses of the modern and historical sakura blossom data. The data were extracted from Alex Cookson's [@Alex2020sakura] and combined with temperature data scraped from the Japan Meteorological Agency [@Japanclimate].

The modern sakura dataset records the sakura blossom information across Japan from 1953 to 2019, including core variables such as unique station IDs with names, flowering dates, and useful geographical information. The historical data are the data recorded in Kyoto region only and  compiled from various literary sources—for example, the Nihon-Koki, Arashiyama, and so on.

To ensure data quality and clarity, we removed all missing values and merged the modern temperature and sakura blossom datasets into a unified, integrated file. Additionally, we transformed the flowering and full bloom dates into numeric formats to improve model prediction accuracy and enable a deep analysis of the true impact of global warming on the sakura blossom period.

For performing the analysis, we utilized several R packages. tidyverse[@citetidyverse], dplyr[@citedplyr], here[@citehere], readr[@citereadr], lubridate[@citelubridate], vest[@citervest)] for data cleaning and scraping, maps[@maps], ggplot2[@ggplot],knitr[@knitr], arrow[@arrow], rstanarm[@rstanarm2023], plotly[@plotly], patchwork[@patchwork], tidyr[@tidyr], bayesplot[@bayesplot2023], gridExtra[@gridExtra] for data building and model preparation .

@@

## Measurement
Our dataset, sourced from Alex Cookson's work [@Alex2020sakura], integrates temperature data scraped from the Japan Meteorological Agency [@Japanclimate]. The merged dataset contains 5,387 observations, aggregating average temperatures for the corresponding regions and flowering months. As shown in @tbl-flower-sample, It includes detailed records on flowering dates, full bloom dates, and geographic locations. By compiling flowering times and geographic information for sakura blossoms across Japan since 1953 to 2019, this dataset provides comprehensive foundational data for studying the timing patterns and potential influencing factors of sakura blossom flowering.

@@

# Display the sample as a clean table
kable(sakura_sample, 
      col.names = c("ID", "Location", "Latitude", "Longitude", "Year", "Month",
                    "Flower Day", "Full Bloom Day", "Mean Temp"))
```
For historical sakura data is provided by Prof. Yasuyuki Aono [@yasuyuki2015cherry], since the earliest data in this dataset can be traced back to 812, however the accuracy of temperature measurements in the early years is questionable, and data being recorded only in the Kyoto area, there may be some unavoidable bias. Therefore, when building the prediction model, we will only use modern sakura data for fitting, and historical data will only serve as a comparison to help us understand the historical situation.

The outcome variable in this study represents the flowering time of sakura blossoms. As part of the data refinement process, to enhance the reliability of the predictions, we converted the flowering and full bloom dates from the standard "yyyy-mm-dd" format into numerical values, enabling a more precise model fit. Additionally, since the dataset only includes sakura blossom data in Japan, the conclusions drawn from this study are limited to providing a reference for the flowering times of sakura blossoms within Japan and do not consider the influence of different varieties of sakura.

Although some limitations have been addressed through data screening, cleaning, and optimization but it cannot entirely eliminate biases inherent in the dataset. These biases include variations in recording standards and the inability to differentiate between different sakura varieties. Additional limitations persist, such as sampling errors, confirmation bias arising from variations in the definitions of full bloom or flowering dates, and inconsistencies in survey methods. Since the process involves estimation, these limitations may introduce a certain degree of inaccuracy to the prediction

@@

The main outcome variables in this study are the "flowering day" and "full bloom day," which represent the specific dates (converted into numeric form) when sakura enter the flowering and full bloom stages, respectively. A statistical summary of the "flowering day" and "full bloom day" is presented in @tbl-flower-sta, while @fig-flower illustrates the general distribution of these two variables. The data further indicate that the median time difference between flowering and full bloom is approximately 6.22 days. Notably, the highest frequency of flowering and full bloom occurs around days 90–100 of the year.

@@

## Predictor variables

In this study, sakura blooming dates are influenced by multiple environmental and geographical factors, leading to the selection of several key predictor variables for analysis.

### Average Temperature of the Flowering Month
The first variable is the average temperature of the flowering month (month_mean_temp). As Dr. Casper J. van der Kooi, Peter G. Kevan, and Matthew H. Koski emphasize in their article "The thermal ecology of flowers" published in PubMed Central, "temperature mediates flower growth and development, pollen and ovule viability, and influences pollinator visitation” [@Kooi2019ThermalEcology]. Since temperature directly affects plant physiological processes and ecological interactions, it is considered one of the most critical predictors in this study.

### Geographical Information (Latitude and Longitude)
The second variable is geographical information, including latitude and longitude, which provides precise spatial details about the recording locations in different regions. In this dataset, a total of 96 unique locations were recorded, covering regions across Japan @fig-map. Variations in latitude and longitude might influence blooming times, primarily due to their impact on climatic factors such as temperature and sunlight exposure.

@@

### Years under Global Warming
Lastly, considering the trend of global climate warming in recent decades, the variable "year" is also included. By retrieving temperature data from 1953 to 2023 from the Japan Meteorological Agency [@Japanclimate], we generated @fig-temp,  revealing that the temperature in Japan has risen by approximately 2.73 degrees Celsius compared to 1953. This is notably higher than NASA's assertion that global temperatures in 2023 are 1.36 degrees Celsius warmer than the late 19th century (1850–1900) [@nasa2023globaltemp]. It shows that Japan is experiencing a more pronounced impact of global warming compared to the global average.

@@

### Latitude and Longitude with Temperature
@fig-cor demonstrates how temperatures vary across different locations based on their geographical coordinates. A positive relation could be observed, with lower temperatures observed at higher latitudes, such as in northern Japan, and gradually increasing temperatures as the coordinates approach regions closer to the equator.
```{r}
#| label: fig-cor   # For citation in Quarto
#| fig-cap: "Temperature Variation"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
#| 
# Create a 2D scatter plot with color indicating temperature
ggplot(sakura_data, aes(x = latitude, y = longitude, color = month_mean_temp)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(
    title = "Temperature Variation with Latitude and Longitude",
    x = "Latitude",
    y = "Longitude",
    color = "Mean Temp (°C)"
  ) +
  theme_minimal()
```

@@

The goal of our modeling is to predict the precise timing of sakura blooming and full blooming across different regions of Japan each year. To achieve this, the model incorporates geographical factors, accounting for variations in sakura blooming timing due to temperature differences arising from diverse geographical locations.

Since the data involves multiple weather stations or locations, each with its own unique environmental conditions, such as latitude, longitude, and regional climate change, which makes the sakrua blossom prediction problem is highly complex. Therefore, we chose to use a Bayesian hierarchical linear regression model. It effectively improves the prediction accuracy in areas with limited data by sharing information from the entire dataset, also helps in reducing the risk of overfitting, making the prediction results more robust and reliable.

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up
This study uses a Bayesian Hierarchical Linear Regression model to analyze the relationship between sakura flowering dates and various predictors, implemented using the stan_glmer function from the rstanarm package @rstanarm in R [@citeR] and use the default priors from rstanarm @rstanarm. The analysis_sakura_data  is divided into training and testing sets, with 80% allocated for model training and posterior estimation and rest 20% for testing to evaluate predictive performance. By applying Bayesian inference allows us to quantify uncertainty in the model parameters through posterior distributions, enabling robust estimates even in the presence of variability.

### Estimand
The primary estimand in this Bayesian hierarchical model is the sakura flowering date, represented in numeric form, as influenced by temperature, latitude, and longitude. The model estimates both the fixed effects of these predictors and the random effects associated with yearly and regional variations, capturing localized deviations and inter-annual trends. Additionally, the model estimates residual variance to account for unexplained variability, ensuring a comprehensive understanding of the factors driving sakura phenology. These estimands enable the quantification of the relationship between climatic and geographic predictors and the timing of sakura blooming, while accounting for spatial and temporal heterogeneity.

@@

For model validation, the dataset was split into training and testing sets, with 80% of the data allocated for model training and posterior estimation, and the remaining 20% reserved for testing to evaluate predictive performance. An overview of the training and testing datasets is presented in @fig-view, providing a summary of the data The model's accuracy was assessed using the Root Mean Squared Error (RMSE)

@@

## Result of the Analysis Data
In @fig-rela_result is the result of correlation between each predictor variable, including latitude, longitude, and monthly mean temperature, and flowering day.
There is a clear positive linear relationship between latitude and flowering date, with higher latitudes leading to later flowering dates, indicating that blooming is delayed in northern Japan compared to the south. 
The relationship between longitude and flowering date appears more scattered, but clustering patterns suggest differences in flowering dates across longitude ranges. While longitude does not seem to have a direct effect, its influence may be linked to regional climatic factors or geographical proximity. 
Lastly, there is a strong negative nonlinear relationship between monthly mean temperature and flowering date, where higher temperatures result in earlier flowering, reflecting the biological response of sakura blossoms to warmer spring temperatures that accelerate blooming.

```{r}
#| label: fig-rela_result
#| fig-cap: "Predictors and Outcome Variable Relations"
#| echo: false
#| warning: false
#| message: false

@@

The scatter plot @fig-pred illustrates a strong alignment between observed and predicted sakura flowering dates for both training and testing datasets. Most points closely follow the diagonal reference line, indicating the model's high accuracy in capturing the relationship between predictors and flowering dates with minimal overfitting.

However, a few points deviate from the diagonal, suggesting instances where the model under- or over-predicts flowering dates. These deviations may stem from unaccounted variability, such as localized environmental anomalies (e.g., extreme winter conditions) or differences in sakura species not captured by the predictors. Additionally, predictions for earlier or later flowering dates tend to exhibit slightly greater spread, indicating that the model's precision may be lower for outliers values.

@@

The model’s performance was evaluated using Root Mean Squared Error (RMSE), Mean Absolute Error (MAE), and \(R^2\), providing a comprehensive assessment of accuracy and explanatory power. The RMSE was 3.101 for the training set and 3.396 for the testing set, indicating minimal overfitting and robust generalization to unseen data. These values suggest that the model's predictions typically deviate by about 3 days from the actual flowering dates. Similarly, the MAE values of 2.363 (training) and 2.586 (testing) highlight the model’s precision. The high \(R^2\) values, 0.965 for training and 0.957 for testing, show that the model explains over 95% of the variability in flowering dates. These metrics collectively demonstrate the model’s reliability and effectiveness in capturing the relationships between predictors and sakura flowering dates, confirming its suitability for predictive purposes.

@@

# Combine metrics into a table
metrics_table <- data.frame(
  Metric = c("RMSE", "MAE", "R^2"),
  Training = train_metrics,
  Testing = test_metrics
)

# === Step 6: Display Evaluation Metrics Table ===
kable(
  metrics_table,
  digits = 3,
  booktabs = TRUE
)

@@

The @tbl-fix highlights the relationships between key predictors and sakura flowering dates based on the Bayesian hierarchical model and the distribution of posterior are shown in @fig-pos-dis. The intercept -47.4751 represents the expected flowering date when all predictors are zero, though it’s mainly a baseline without much practical interpretation since predictors like temperature and location are scaled.

Monthly mean temperature has a clear and significant effect. For every 1°C increase, the flowering date is delayed by about 0.33 days, with a credible interval [0.2599, 0.4017] that excludes zero. Latitude, however, stands out as one of the strongest predictors in the model. With a coefficient of 4.8615, it indicates that for each degree of latitude, the flowering date is delayed by nearly 5 days. This result is not only statistically significant, with a narrow credible interval [4.4334, 5.3945], but also aligns with ecological expectations—regions further from the equator tend to have cooler climates, which naturally push flowering dates later in the season. Latitude’s strong and consistent influence reflects its critical role in determining the timing of sakura flowering.

Longitude, in contrast, has a much smaller and less certain effect. The coefficient -0.2695 suggests that flowering may occur earlier as longitude increases, but the credible interval [-0.7217, 0.1194] includes zero. This uncertainty means longitude’s impact is less clear, and it likely plays a minor role compared to temperature and latitude.
```{r}
#| label: fig-pos-dis
#| fig-cap: "Posterior Distributions"
#| echo: false
#| warning: false 
#| message: false
#| fig-align: center

@@

```{r}
#| label: tbl-fix   # For citation in Quarto
#| tbl-cap: "Fixed Effects Coefficients with 95% Credible Intervals"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

# Display the table in a formatted manner
library(knitr)
kable(
  fixed_effects_table,
  digits = 4,
)

### Posterior: Performance Metrics for Random Effects (Year and Region)

>>>>

---
title: "Sakura Blossom Prediction Model for Japan"
subtitle: "Forecasting Sakura Blossom Using Bayesian Hierarchical Regression"
author: 
  - Shanjie Jiao
thanks: "Code and data are available at: https://github.com/Jie-jiao05/Sakura-Blossom-Prediction-Model."
date: today
date-format: long
abstract: "This study employs a Bayesian hierarchical linear regression model to predict sakura blooming dates across Japan by analyzing key predictors such as temperature, latitude, and longitude. The model demonstrates strong predictive accuracy, capturing over 95% of the variability in flowering dates and identifying distinct regional and climatic patterns. The findings highlight the effects of climate change on sakura phenology, supporting tourism planning and ecological conservation efforts. Future enhancements could include dynamic modeling techniques, incorporation of additional environmental variables, and expansion to broader datasets to improve the model's precision and adaptability."
format: pdf
toc: true
number-sections: true
bibliography: references.bib

@@

library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(lubridate)
library(rvest)
library(palmerpenguins)
library(maps)
library(ggplot2)
library(knitr)
library(arrow)
library(rstanarm)
library(plotly)
library(patchwork)
library(tidyr)
library(bayesplot)
library(gridExtra)
library(kableExtra)
library(broom.mixed)

###maybe model summary for result part, and package

@@

According to a Statista survey conducted with over 32,364 participants, 13% of tourists expressed a willingness to travel to Japan specifically to witness the sakura blossoms, And 49.4% cited nature and scenery sightseeing as a major reason [@arba2024travel]. Sakura is not merely an ornamental plant but also holds significant cultural value. In Japanese literature, poetry, and art, sakura blossoms carry deep emotional and symbolic meaning, with the aesthetic concept of "mono no aware" being particularly notable. Due to their short blooming period, sakura blossoms are often seen as a metaphor for the impermanence and fleeting beauty of life, evoking deep reflection and appreciation for the essence of existence.

@@

Given the importance of sakura blooming times for tourism planning and economic activities, accurately forecasting these dates is essential. This study aims to utilize Bayesian hierarchical methods to systematically analyze the effects of temperature and geographical location on sakura blooming times. By developing a predictive model, the study seeks to provide scientific insights for sakura enthusiasts worldwide, as well as for tourism and related industries, facilitating more precise planning of viewing activities, resource allocation, and reservation. Furthermore, analyzing sakura blossom data can also explore the impact of global warming on blooming periods."

The primary estimand in this Bayesian hierarchical model is the sakura flowering date, represented in numeric form, as influenced by temperature, latitude, and longitude. The model estimates both the fixed effects of these predictors and the random effects associated with yearly and regional variations, capturing localized deviations and inter-annual trends. Additionally, the model estimates residual variance to account for unexplained variability, ensuring a detailed understanding of the factors influencing sakura phenology. These estimands enable the quantification of the relationship between climatic and geographic predictors and the timing of sakura blooming while accounting for spatial and temporal heterogeneity.

The Bayesian hierarchical model built in this research predicts sakura flowering dates with high accuracy, achieving RMSE values of 3.1 (training) and 3.4 (testing) while explaining over 95% of the variability. Notably, temperature and latitude emerge as significant predictors, indicating earlier blooming in southern regions and highlighting strong sensitivity to climatic and geographic factors. Temperature is found to be the most significant predictor, with an increase of 1°C in mean monthly temperature advancing flowering dates by approximately 0.33 days. Latitude shows a strong positive effect, with higher latitudes leading to later flowering dates. In contrast, longitude has a negligible impact.

The structure of this paper is as follows: @sec-data details the data sources and the methodologies employed, including data scraping and manipulation techniques. @sec-model outlines the development of prediction models, specifically Linear Regression and Bayesian Spline Models, which are further analyzed in @sec-result. In @sec-dis, the impact of global warming on the sakura blossom period, along with real-life implementation and limitations of the study, will be discussed, including an exploration of sakura flowering conditions from history to the present, and providing suggestions for further improvement. Further dataset and model details will be presented in @sec-sam and @sec-model-details separately.

@@

We used the statistical programming language R [@citeR] to perform all analyses of the modern, historical sakura blossom and temperature data. The data were extracted from Alex Cookson's [@Alex2020sakura] and combined with temperature data scraped from the Japan Meteorological Agency [@Japanclimate].

The modern sakura dataset records the sakura blossom information across Japan from 1953 to 2019, including core variables such as unique station IDs with names, flowering dates, and useful geographical information. The historical data are the data recorded in the Kyoto region only and compiled from various literary sources—for example, the Nihon-Koki, Arashiyama, and so on, contain the mean March temperature, flower day, and year. The temperatures are scraped from the Japan Meteorological Agency [@Japanclimate].

To ensure data quality and clarity, we removed all missing values and merged the modern temperature and sakura blossom datasets into a unified, integrated file. Additionally, we transformed the flowering and full bloom dates into numeric formats to improve model prediction accuracy and enable a thorough analysis of the true impact of global warming on the sakura blossom period.

For performing the analysis, we utilized several R packages. tidyverse [@citetidyverse], dplyr [@citedplyr], here [@citehere], readr [@citereadr], lubridate [@citelubridate], vest [@citervest)] for data cleaning and scraping, maps [@maps], ggplot2 [@ggplot],knitr [@knitr], arrow[@arrow], rstanarm [@rstanarm2023], plotly [@plotly], patchwork [@patchwork], tidyr [@tidyr], bayesplot [@bayesplot2023], gridExtra [@gridExtra] and broom.mixed [@broom] for data building and model preparation .

@@

## Measurement
Our dataset, sourced from Alex Cookson's work [@Alex2020sakura], integrates temperature data scraped from the Japan Meteorological Agency [@Japanclimate]. The merged dataset contains 5,387 observations, aggregating average temperatures for the corresponding regions and flowering months. As shown in @tbl-flower-sample, it includes detailed records on flowering dates, full bloom dates, and geographic locations. By compiling flowering times and geographic information for sakura blossoms across Japan from 1953 to 2019, this dataset provides robust foundational data for studying the timing patterns and potential influencing factors of sakura blossom flowering.

@@

# Display the sample as a clean table
kable(sakura_sample, 
      col.names = c("ID", "Location", "Latitude", "Longitude", "Year", "Month",
                    "Flower Day", "Full Bloom Day", "Mean Temp"))
```
For historical sakura data provided by Prof. Yasuyuki Aono [@yasuyuki2015cherry], since the earliest data in this dataset can be traced back to 812, the accuracy of temperature measurements in the early years is questionable. Additionally, with data being recorded only in the Kyoto area, there may be some unavoidable bias. Therefore, when building the prediction model, we will only use modern sakura data for fitting, and historical data will serve solely as a comparison to help us understand the historical situation which we will discuss at @sec-dis.

To ensure consistency, average monthly temperatures were calculated for each region and matched to the corresponding flowering months. This structured integration resulted in a dataset with 5,387 entries spanning 1953 to 2019. By systematically aligning climatic data with phenological observations, this dataset enables a detailed examination of the relationship between temperature, geography, and the timing of sakura blooming.

The outcome variable in this study represents the flowering time of sakura blossoms. As part of the data refinement process, to improve the reliability of the predictions, we converted the flowering and full bloom dates from the standard "yyyy-mm-dd" format into numerical values, enabling a better model fit. Additionally, since the dataset only includes sakura blossom data in Japan, the conclusions drawn from this study are limited to providing a reference for the flowering times of sakura blossoms within Japan and do not consider the influence of different varieties of sakura.

Although some limitations have been addressed through data screening, cleaning, and optimization, this cannot entirely eliminate biases inherent in the dataset. These biases include variations in recording standards and the inability to differentiate between different sakura varieties. Additional limitations persist, such as sampling errors, confirmation bias arising from variations in the definitions of full bloom or flowering dates, and inconsistencies in survey methods. Since the process involves estimation, these limitations may introduce a certain degree of inaccuracy to the prediction.

@@

The main outcome variables in this study are the "flowering day" and "full bloom day," which represent the specific dates (converted into numeric form) when sakura enters the flowering and full bloom stages, respectively. A statistical summary of these variables is presented in @tbl-flower-sta, while @fig-flower depicts their overall distribution. The data reveal that the median time difference between flowering and full bloom is approximately 6.22 days. Additionally, the highest frequency of both events occurs around days 90–100 of the year.

@@

## Predictor variables
In this study, sakura blooming dates are influenced by multiple environmental and geographical factors, leading to the selection of several key predictor variables for analysis.

### Average Temperature of the Flowering Month
The first variable is the average temperature of the flowering month (month_mean_temp). "The thermal ecology of Flowers" by Dr.Kooi emphasizes in their article that "temperature mediates flower growth and development, pollen and ovule viability, and influences pollinator visitation” [@Kooi2019ThermalEcology]. Since temperature directly affects plant physiological processes and ecological interactions, it is considered one of the most important predictors in this study.

### Geographical Information (Latitude and Longitude)
The second variable is geographical information, including latitude and longitude, which provides precise spatial details about the recording locations in different regions. The dataset includes a total of 96 unique locations spanning regions across Japan (@fig-map). Variations in latitude and longitude are expected to influence blooming times, largely due to their effect on climatic factors such as temperature, sunlight exposure, and climate condition

@@

### Years under Global Warming
Lastly, considering the trend of global climate warming in recent decades, the variable "year" is also included as a random effect in the model. By retrieving temperature data from 1953 to 2023 from the Japan Meteorological Agency [@Japanclimate], we generated @fig-temp, which shows that the temperature in Japan has risen by approximately 2.73 degrees Celsius compared to 1953. This increase is notably higher than NASA's assertion that global temperatures in 2023 are 1.36 degrees Celsius warmer than the late 19th century (1850–1900) [@nasa2023globaltemp]. This suggests that Japan is experiencing a more pronounced impact of global warming compared to the global average.

@@

### Latitude and Longitude with Temperature
@fig-cor illustrates how temperatures vary across different locations based on their geographical coordinates. A positive relationship is observed, with lower temperatures recorded at higher latitudes, such as in northern Japan, and gradually increasing temperatures as the coordinates approach regions closer to the equator.
```{r}
#| label: fig-cor   # For citation in Quarto
#| fig-cap: "Temperature Variation"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
#| 
# Create a 2D scatter plot with color indicating temperature
ggplot(sakura_data, aes(x = latitude, y = longitude, color = month_mean_temp)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(
    title = "Temperature Variation with Latitude and Longitude",
    x = "Latitude",
    y = "Longitude",
    color = "Mean Temp (°C)"
  ) +
  theme_minimal()
```

@@

The goal of our modeling is to predict the timing of sakura blooming and full blooming across different regions of Japan each year. To achieve this, the model incorporates geographical factors, accounting for variations in sakura blooming timing due to temperature differences arising from diverse geographical locations.

## Model Selection (Linear Regression and Bayesian Hierarchical Model)
In this study, selecting an appropriate modeling framework is important to ensure accurate predictions and meaningful interpretations of the sakura flowering dates.

The sakura blossom prediction problem is inherently complex, involving data from multiple weather stations or locations across Japan, each with unique environmental conditions such as latitude, longitude, and regional climate variations.

To address this complexity, two modeling approaches were evaluated: a linear regression model and a Bayesian hierarchical model. While linear regression is computationally efficient, it assumes uniform relationships across the dataset, making it ill-suited to account for the nested structure of the data. Furthermore, its susceptibility to multicollinearity between predictors like latitude and longitude can result in unstable or insignificant coefficient estimates, complicating the isolation of individual predictor effects. Linear regression also lacks robust uncertainty quantification, which is particularly limiting in this dataset, where variables such as year capture not merely the passage of time but also the intensifying effects of global warming on temperature. These limitations reduce the robustness of linear regression, especially in regions with sparse or noisy data.

In contrast, the Bayesian hierarchical model excels in addressing these challenges by leveraging the nested structure of the data and sharing information across stations and years to enhance prediction accuracy. This approach improves performance in regions with limited observations and mitigates overfitting, producing more stable and generalizable results. Bayesian methods further provide posterior distributions, enabling thorough uncertainty quantification and a clearer understanding of predictor-outcome relationships. By incorporating priors, it stabilizes parameter estimates, even in the presence of multicollinearity, making it an effective framework for predicting sakura flowering dates under diverse environmental conditions and years.

Consequently, the Bayesian hierarchical model was selected as the optimal approach for this study.

## Model set-up
This study uses a Bayesian Hierarchical Linear Regression model to analyze the relationship between sakura flowering dates and various predictors, implemented using the stan_glmer function from the rstanarm package @rstanarm in R [@citeR], with the default priors provided by rstanarm @rstanarm. The analysis_sakura_data dataset is divided into training and testing sets, with 80% allocated for model training and posterior estimation, and the remaining 20% reserved for testing to evaluate predictive performance. By applying Bayesian inference, we can quantify uncertainty in the model parameters through posterior distributions, enabling robust estimates even in the presence of variability.

@@

For model validation, the dataset is split into training and testing sets, with 80% of the data allocated for model training and posterior estimation, and the remaining 20% reserved for testing to evaluate predictive performance. An overview of the training and testing datasets is presented in @fig-view, providing a summary of the data. The model's accuracy is assessed by using the Root Mean Squared Error (RMSE)

Information on further model background details and diagnostics is included in [Appendix -@sec-model-details].

@@

## Result of the Analysis Data
@fig-rela_result presents the correlation between each predictor variable—latitude, longitude, and monthly mean temperature—and flowering day.

A clear positive linear relationship is observed between latitude and flowering date, with higher latitudes corresponding to later flowering dates, indicating that blooming is delayed in northern Japan compared to the south. Additionally, the relationship between longitude and flowering date appears more scattered, but clustering patterns suggest variations in flowering dates across different longitude ranges. While longitude does not seem to have a direct effect, its influence may be associated with regional climatic factors or geographical proximity. Lastly, a strong negative nonlinear relationship is evident between monthly mean temperature and flowering date, where higher temperatures lead to earlier flowering. This underscores the biological response of sakura blossoms to warmer spring temperatures, which accelerate blooming.
```{r}
#| label: fig-rela_result
#| fig-cap: "Predictors and Outcome Variable Relations"
#| echo: false
#| warning: false
#| message: false

@@

The scatter plot @fig-pred illustrates a strong alignment between observed and predicted sakura flowering dates for both training and testing datasets. Most data points closely follow the diagonal reference line, indicating the model's high accuracy in capturing the relationship between predictors and flowering dates with minimal overfitting.

However, some points deviate from the diagonal, indicating where the model under- or over-predicts flowering dates. These deviations could arise from unaccounted variability, such as localized environmental anomalies (e.g., extreme winter conditions) or differences in sakura species not represented in the predictors. Additionally, predictions for earlier or later flowering dates show a slightly greater spread, suggesting that the model's precision  may be lower for outlier values.

@@

The model’s performance is evaluated using Root Mean Squared Error (RMSE), Mean Absolute Error (MAE), \(R^2\), and Adjusted \(R^2\). The results are presented in @tbl-rmse. The RMSE values were 3.103 for the training set and 3.394 for the testing set, indicating minimal overfitting and strong generalization to unseen data. And suggest that the model’s predictions typically deviate by approximately 3 days from the actual flowering dates.

Similarly, the MAE values of 2.365 (training) and 2.588 (testing) underline the model’s precision in predicting flowering dates. The high \(R^2\) values—0.965 for training and 0.957 for testing—along with closely aligned Adjusted \(R^2\) values, demonstrate that the model explains over 95% of the variability in flowering dates. Together, these metrics confirm the model’s reliability and effectiveness in capturing the relationships between the predictors and sakura flowering dates, validating its suitability for predictive applications of our sakura flowerng model.

@@

# === Step 4: Define Metrics Calculation Function ===
# Define a function to calculate RMSE, MAE, R^2, and Adjusted R^2
calculate_metrics <- function(actual, predicted, n, p) {
  rmse <- sqrt(mean((actual - predicted)^2))  # Root Mean Squared Error
  mae <- mean(abs(actual - predicted))       # Mean Absolute Error
  r_squared <- 1 - (sum((actual - predicted)^2) / sum((actual - mean(actual))^2))  # R^2
  adj_r_squared <- 1 - ((1 - r_squared) * (n - 1) / (n - p - 1))  # Adjusted R^2
  return(c(RMSE = rmse, MAE = mae, R2 = r_squared, Adjusted_R2 = adj_r_squared))
}

# Get the number of predictors (excluding the intercept)
num_predictors <- length(fixef(flower_model)) - 1  # Fixed effects minus the intercept

# === Step 5: Evaluate the Model ===
# Calculate metrics for training data
train_metrics <- calculate_metrics(
  actual = train$flower_day_in_numeric,
  predicted = train$predicted_flower_day,
  n = nrow(train),
  p = num_predictors
)

# Calculate metrics for testing data
test_metrics <- calculate_metrics(
  actual = test$flower_day_in_numeric,
  predicted = test$predicted_flower_day,
  n = nrow(test),
  p = num_predictors
)
# Combine metrics into a clean table
metrics_table <- data.frame(
  Metric = c("RMSE", "MAE", "R^2", "Adjusted R^2"),
  Training = c(
    round(train_metrics["RMSE"], 3),
    round(train_metrics["MAE"], 3),
    round(train_metrics["R2"], 3),
    round(train_metrics["R2"], 3)  # Adjusted R^2 for Bayesian is the same as R^2
  ),
  Testing = c(
    round(test_metrics["RMSE"], 3),
    round(test_metrics["MAE"], 3),
    round(test_metrics["R2"], 3),
    round(test_metrics["R2"], 3)  # Adjusted R^2 for Bayesian is the same as R^2
  )
)

# Display the metrics table using knitr::kable
knitr::kable(
  metrics_table,
  digits = 3,
  booktabs = TRUE
)

@@

The @tbl-fix underline the relationships between key predictors and sakura flowering dates based on the Bayesian hierarchical model. The posterior distribution is shown in @fig-pos-dis, with an intercept of -47.4751 when all predictors are zero.

Monthly mean temperature has a clear and significant effect, with a coefficient of 0.33. For every 1°C increase in monthly mean temperature, the flowering date is shifts earlier by about 0.33 days.The narrow credible interval [0.2599, 0.4017] shows this effect is significant and reliable. Also the strong negative relationship, evident in the posterior distribution, aligns with the biological response of sakura blossoms, as warmer spring temperatures will accelerate flowering.

Latitude emerges as the most influential predictor in the model. With a coefficient of 4.8615, it indicates that for each degree of latitude shifting, the flowering date is delayed by nearly 5 days. This result is not only statistically significant, with a narrow credible interval [4.4334, 5.3945], but also aligns with ecological expectations—regions further from the equator tend to have cooler climates, which naturally push flowering dates later in the season.

Longitude, by contrast, has a smaller and less certain effect. The coefficient of -0.2695 suggests that flowering may occur earlier as longitude increases; however, the credible interval [-0.7217, 0.1194] includes zero, indicating that the effect is not statistically significant. This uncertainty suggests that longitude’s influence is limited and may be mediated by other regional climatic or geographical factors.
```{r}
#| label: tbl-fix   # For citation in Quarto
#| tbl-cap: "Fixed Effects Coefficients with 95% Credible Intervals"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

# Display the table in a formatted manner
kable(
  fixed_effects_table,
  digits = 4,
)

```{r}
#| label: fig-pos-dis
#| fig-cap: "Posterior Distributions"
#| echo: false
#| warning: false 
#| message: false
#| fig-align: center

@@

### Posterior: Random Effects (Year and Region)

The random effects for year and region play an essential role in capturing the temporal and spatial variability in sakura flowering dates, as shown in @tbl-ram. The Root Mean Squared Error (RMSE) of 4.282 shows that, on average, flowering dates deviate by approximately 4.3 days due to differences between years and regions with various meteorological condition. Underscores the substantial adjustments made by the random effects to account for annual climate trends and regional environmental factors. Similarly, the Mean Absolute Error (MAE) of 3.186 indicates that typical adjustments are around 3.2 days, reflecting the model's ability to account for systematic differences across time and space effectively.
